---
title: "R for developers"
author: "Steph Locke"
date: "`r Sys.Date()`"
output: 
  html_document: 
    df_print: tibble
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=TRUE, message = FALSE, warning = FALSE)
```
# Upfront stuff
## Agenda
- Why you should know R
- What are the ways you can write R
- How you can help R users improve their code
- How R users can consume your code
- How you can get R code into production

## Me
- Data Scientist
- Microsoft AI MVP
- Microsoft Machine Learning MCT
- Author: geni.us/rfundamentals
- Locke Data
- t: @theStephLocke

## Locke Data
- Data Science Consultancy
- Focussed on helping orgs *start* 
- Site: itsalocke.com
- t: @lockedata

## This talk
- Get the slides at itsalocke.com/talks
- Grab a sticker and/or a business card
- Book giveaway for good questions
- Limited supply of tshirts available

# Why you should know R
```{r whyyoushouldknowr}
```

## Why
- Data Science
- Data wrangling
- Cross-platform capability
- Reproducibility
- Efficeincy
- Flexibility
- Cash

## Why not Python?
> It isn't one or the other

There's plenty of ways for these two languages to interoperate and co-exist. Pick the right hammer for a given nail!

## Data profiling
```{r eval = FALSE}
DataExplorer::create_report(iris)
```

## Nice webscraping

```{r, results='hide', echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(rvest)
```

```{r tidy=FALSE}
library(tidyverse)
library(rvest)

"https://en.wikipedia.org/wiki/List_of_superhero_debuts" %>%
  read_html() %>%
  html_nodes(xpath = "//*[@id='mw-content-text']/div/table") %>%
  map(html_table, fill = TRUE) %>%
  map_df( ~ mutate(., `Year Debuted` = as.character(`Year Debuted`))) %>%
  mutate(`Char_Team` = coalesce(`Character / Team / Series`, `Character / Team`)) %>%
  select(Char_Team, Year_Debut = `Year Debuted`) %>%
  mutate(Year = str_extract(Year_Debut, "[0-9]{4}")) %>%
  skimr::skim()
```


## Great dataviz
```{r}
library(leaflet)
data(quakes)

# Show first 20 rows from the `quakes` dataset
leaflet(data = quakes[1:20,]) %>% addTiles() %>%
  addMarkers(~long, ~lat, popup = ~as.character(mag), label = ~as.character(mag))
```

## Super apps
[Cost of Coffee](https://stef.shinyapps.io/cost_of_coffee/)

## Data driven content
- R
- Python
- JS
- Bash
- SQL

# Ways you can write R
```{r waysyoucanwriter}
```

## Variables, comments, etc
```{r}
# Define a variable
a<-25

# Call a variable
a

# Do something to it
a+1
```

## Structures
```{r}
# Vector
a = c(25, 30)

# Matrix
matrix(a)

# Data frame
data.frame(a,b=a/5,c=LETTERS[1:2])

# List
list(vector=a, matrix=matrix(a))
```

## Functions
```{r}
# Define a function
showAsPercent<-function(x) {
  paste0(round(x*100 ,0) ,"%")
}

# Call a function
showAsPercent(0.1)
```

## Extending R
```{r, eval=FALSE }
# Get a package
install.packages("tibble")

# Activate a package
library(tibble)

# Use the package as a namespace
tibble::as.tibble(a)
```

## Base (similar to pandas)

```{r}
df <- data.frame(a=1:10, b = LETTERS[1:5]) # setup
df[1:2,] # row numbers
df[df$a<2,] # value filters
```

## Base (similar to pandas)
```{r}
df[df$a<3,1] # column filter
df[df$a<3,1, drop=FALSE] # column filter (keep data.frame)
```


## OO
```{r, results='hide', echo=FALSE, warning=FALSE, message=FALSE}
library("R6")
```

```{r}
# Orig OO (s3): cyclismo.org/tutorial/R/s3Classes.html
library(R6)
Loan<-R6Class("Loan", 
              public=list(term=NA
                         ,initialize=function(term){
                           if(!missing(term)){ 
                              self$term<-term 
                              }} 
                         ,extendBy=function(ext){ 
                            self$term<-self$term+ext
                            }))
```


## OO
```{r}
acc<-Loan$new(36)
acc$extendBy(6)
acc$term
```

## Functional
```{r, results='hide', echo=FALSE, warning=FALSE, message=FALSE}
library(magrittr)
```

```{r}
library(magrittr)
iris %>% 
  {tail(.,nrow(.)/5)} %>% 
  pairs()
```

## Functional

```{r}
library(tidyverse)
iris %>% 
  filter(Petal.Width<2) %>%
  group_by(Species) %>%
  summarise_all(mean)
```


## Best practices
- Make sure to declare packages using `library()` at the top of the script
- Label sections of work using `# ---- SectionName ----` to allow you to pick up the code into a LaTeX or markdown doc later
- Never delete <- always make new objects with modifications
- If comfortable with coding, or have a lot of data to process use `data.table` over `dplyr`
- Try to do all data manipulation at the top of the script
- Reuse your code by writing functions for anything you do frequently



## Documentation
Document as you go!

- Use markdown to use a light syntax for integrating your code and commentary
- Use LaTeX for finer level of control on output layout
- Use these instead of Excel or Word so that a change in assumption means updating the code and re-running the doc
- Load your main R file up and use the labels from `# ---- SectionName ----` to save repetition
- Present in HTML, slide decks, PDF, Word at the click of a button

## Workflow best practices
- Focus on working code & documentation
- Use source control (github)
- Write tests (`assertive`, `assertthat`, `testthat`)
- Regularly visualise (`ggplot2`)
- Do code reviews
- Modularise
- Use Rstudio
- Consider structuring as a package

## Next steps
- geni.us/rfundamentals
- notebooks.azure.com/stephlocke/libraries
- #rstats
- datacamp.com
- r-bloggers.com

# How you can help R users improve their code
```{r helprusersimprovetheircode}
```

## Building packages
github.com/lockedata/pRojects
geni.us/rpkgs

```{r eval=FALSE}
devtools::install_github("lockedata/projects")
pRojects::createPackageProject()
```

## Writing tests
- testthat
- stubthat
- mockr

```{r}
library(testthat)
context("Some thing I want to test")
test_that("things make sense",
          expect_equal(1, 1))
```

## Code coverage
```{r}
library(covr)
?package_coverage
?codecov
```

## Continuous integration
- Travis CI

# How R users can consume your code

## JavaScript
```{r}
library(V8)
ct<-v8()
ct$eval("var foo = 123")
ct$eval("var bar = 456")
ct$eval("foo + bar")

ct$source(system.file("js/underscore.js", package="V8"))
ct$call("_.filter", mtcars, JS("function(x){return x.mpg < 15}"))
```


## Python
```{r}
library(reticulate)
main <- import_main()
builtins <- import_builtins()
builtins$print('foo')
py_run_string("x = 10")
# access the python main module via the 'py' object
py$x
```

## APIs
```{r}
library(httr)
"https://foaas.com/version" %>% 
  GET() %>% 
  content()
```

# How you can get R code into production
```{r rtoproduction}
```

## plumber
rplumber.io

```{r eval=FALSE}
#' Echo the parameter that was sent in
#' @param msg The message to echo back.
#' @get /echo
function(msg=""){
  list(msg = paste0("The message is: '", msg, "'"))
}
```

## opencpu
- opencpu.org
- jsfiddle.net/user/opencpu/fiddles/

## docker 
- containerit
- rize
- liftr

## Hadoop(ish)
- sparklyr
- ML Server
- Jupyter

## Microsoft
- SQL Server sp_execute_external_script()
- ML server 
- Azure ML Studio
- Power BI

## Shiny
- shinyserver
- shinyapps.io
- shinyproxy.io

# Thank you!
## Covered
- Why you should know R
- What are the ways you can write R
- How you can help R users improve their code
- How R users can consume your code
- How you can get R code into production

## Follow up
- itsalocke.com/talks
- geni.us/rfundamentals
- t: @theStephLocke
- t: @LockeData
