---
title: 'R for reporting'
author: "Steph Locke\n@SteffLocke\nLocke Data"
date: '`r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=TRUE, size="large", tidy.opts=list( width.cutoff=60))
```

```{r autodoc, child='AboutMeTech.Rmd', eval=TRUE}
```

## Agenda
- R for interactive reports
- R for static reports
- Architectures
- Skills
- Other considerations

# R for interactive reports

## shiny
- inputs
- reactive intermediates
- outputs

## shiny++
- shinydashboard
- flexdashboard

# R for static reports

## rmarkdown
- PDF
- Word
- HTML

## Excel
- writexl
- openxlsx

## Scheduling & email
- [taskscheduleR](https://CRAN.R-project.org/package=taskscheduleR)
- [cronR](https://cran.r-project.org/package=cronR)
- [mailR](https://cran.r-project.org/package=mailR)
- [ponyexpress](https://github.com/ropenscilabs/ponyexpress)


# Architectures

## shiny
- shinyserver
- shinyapps.io

## general
- rstudio connect
- AN Other hosting
- docker

# Skills

## R
- Basic R
- Data manipulation
- Data visualisation
- Writing your own functions and packages
- Shiny

## Others
- Web (HTML / CSS / JS)
- typesetting (LaTeX)
- infrastructure
- dev best practices

# Other considerations

## Other considerations
- Team practices
- Deployments
- Migration
- Dependency management

# Conclusion

## Agenda
- R for interactive reports
- R for static reports
- Architectures
- Skills
- Other considerations

## Wrapup
- Thank you
- Get the slides via [itsalocke.com](https://itsalocke.com/talks)
- Start learning R with [Working with R](http://geni.us/workingwithr)
- Get in touch! [\@SteffLocke](https://twitter.com/stefflocke)
