---
title: 'Intro to R'
author: "Steph Locke\n@SteffLocke\nLocke Data"
date: '`r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=TRUE, size="large", tidy.opts=list( width.cutoff=60))
```

# Setup

## Steph Locke {data-background="https://itsalocke.com/img/ChibiStephBusiness.svg"}
- Locke Data
- Presenter on Data Science, BI, and DataOps
- Microsoft Data Platform MVP
- User group leader: [MSFT Stack](http://www.meetup.com/MSFT Stack) & [CaRdiff](http://www.meetup.com/Cardiff-R-User-Group/)
- Conference organiser: [satRdays](http://planning.satrdays.org)
- Author: [R Fundamentals](//geni.us/rfundamentals)

[github.com/StephLocke](https://github.com/StephLocke) | [itsalocke.com](http://itsalocke.com) | [\@SteffLocke](https://twitter.com/stefflocke)


## Agenda
- What is R?
- R for data manipulation
- R for data visualisation
- R for reporting
- ~R for data science~
- Nifty stuff
- Next steps

# R

## R 
- What is it?
- Origins
- Quirks
- Value

## Use case - wrangling data
```{r}
library(tidyverse,quietly = TRUE)
head(population, 2)
head(spread(population, year, population), 2)
```

## Use case - visualisation
```{r}
library(datasauRus)
ggplot(datasaurus_dozen, aes(x,y)) +
  geom_point() +
  facet_wrap(~dataset, ncol=5)
```

## Use case - integration
```{r}
library(reticulate)
os <- import("os")
os$getcwd()
```

# R for data manipulation

## Tidyverse
- Lots of data sources and destinations
- SQL-ish functions
- Kicks SQLs butt 
- "Piping" for easy to read code

## I/O packages
- readr for flat files
- readxl for Excel
- haven for SAS, Stata, etc
- odbc for databases
- jsonlite for JSON
- xml2 for XML
- httr for APIs
- rvest for web scraping
- sparklyr for Spark

## Data manipulation packages
- dplyr for "SQL"
- dbplyr for "SQL" against DBs
- tidyr for pivoting and unpivoting data
- purrr for loop(ish) stuff without the loops
- magrittr for pipelines

## Bulk read example
```{r}
list.files("../demo/multiread", full.names = TRUE) %>% 
  map_df(read_csv, 
         col_types=cols(.default = col_character()))
```

## Data pipelines
```{r}
starwars %>% 
  unite(homeworld_species, homeworld, species) %>% 
  count(homeworld_species)
```

## Databases
```{r echo=TRUE}
library(DBI)
library(odbc)

driver   = "ODBC Driver 13 for SQL Server"
server = "mhknbn2kdz.database.windows.net"
database = "AdventureWorks2012"
uid = "sqlfamily"
pwd = "sqlf@m1ly"

dbConn<-dbConnect(odbc(),
          driver=driver, server=server,
          database=database, uid=uid,
          pwd=pwd)
```

## Querying databases
```{r echo=TRUE}
library(dbplyr)
 
medalist<-tbl(dbConn, in_schema("olympics","medalist"))
sport<-tbl(dbConn, in_schema("olympics","sport"))

medalist %>% 
  left_join(sport)
```

# R for data visualisation

## Key packages
- ggplot2 for static viz
    + ggthemes for extra themes
    + ggraph for graph viz in ggplot
    + ggmap for geospatial in ggplot
- plotly for interactive viz
- htmlwidgets for generating anything html / javascript
- DiagrammeR for flowcharts and things

## ggplot2 basics
```{r}
ggplot(datasaurus_dozen, aes(x,y)) +
  geom_point() +
  facet_wrap(~dataset, ncol=5)
```

## Turning a ggplot interactive
```{r}
library(plotly, quietly = TRUE)
p<- ggplot(datasaurus_dozen, aes(x,y)) +
  geom_point() +
  facet_wrap(~dataset, ncol=5)
ggplotly(p)
```

## Building interactive plots
```{r}
map_data("world", "UK") %>%
  group_by(group) %>%
  plot_geo(x = ~long, y = ~lat) %>%
  add_markers(size = I(1))
```

# R for reporting
## Key packages
- rmarkdown for using markdown
- knitr for converting R markdown files to other formats
- revealjs for whizzy presentations
- bookdown for making books
- blogdown for sites
- officer for complex(ish) Word and PowerPoint
- shiny for interactive reports
- flexdashboard for dashboards
- DataExplorer for quick reports on data

## rmarkdown & revealjs
See slide source code!

## bookdown
See Data Manipulation in R source code

## shiny & flexdashboard
See Cost of Coffee app & source code

## DataExplorer
```{r eval=FALSE}
DataExplorer::create_report(who)
```


# Nifty stuff
## Plays well with others
- Consume Python & JS
- Consume from Python & JS
- Incoporated into many BI platforms
- Run in-database
- Run in Spark

## Community
- R UGs
- R Ladies
- Conferences e.g. SatRday Cardiff June 23rd
- Twitter (#rstats)
- r-bloggers.com
- r-weekly.com

# Next steps
## Q&A
## Learn
- Community
- datacamp.com
- <shameless plug>My books</shameless plug>

## Follow up with me
- @stefflocke
- Join my #learnr slack! bit.ly/ldlearnrslack
- Get slides and blog posts itsalocke.com
